% ============================================================
%  GENERAL INTRODUCTION
%  File    : divers/introduction.tex
%  Author  : Zenfour Abdelmouaiz  —  ENSTICP 5th Year
%  Thesis  : Zero-Touch Deployment and Cloud-Native
%             Orchestration of Edge Nodes for IoT
%  Partner : AET Industries
%  Rules   : ENSTICP — impersonal tone, deductive paragraphs,
%            no floats, ≤ 2 pages, IEEE citations [n],
%            1 cm first-line indent, justified alignment
% ============================================================

\chapter*{General Introduction}
\addcontentsline{toc}{chapter}{General Introduction}
\markboth{General Introduction}{General Introduction}

% ------------------------------------------------------------------
%  PARAGRAPH 1 — General context: the IoT scaling challenge
% ------------------------------------------------------------------
    The global proliferation of Internet of Things (\textit{IoT}) devices is
reshaping the operational requirements of embedded systems across industrial
sectors. Manufacturing plants, energy installations, and Oil \& Gas facilities
now deploy large fleets of heterogeneous, geographically distributed hardware
nodes, each requiring continuous software management, configuration updates,
and workload lifecycle operations throughout its service lifetime. As fleet
sizes scale from tens to thousands of units, the classical maintenance model —
in which a field technician physically visits each deployment site to perform
firmware updates or reconfigurations — becomes prohibitively costly and
impossible to sustain at scale. This dependency on physical human intervention,
widely referred to in the industry as the \textit{truck roll} problem,
constitutes a fundamental barrier to the scalable operation of modern
IoT infrastructure.

% ------------------------------------------------------------------
%  PARAGRAPH 2 — Specific context and partner: AETI's challenge
% ------------------------------------------------------------------
    \textit{AET Industries} (\textit{AETI}), an Algerian company specialising
in the full lifecycle of embedded hardware innovation — from concept and firmware
development through to prototyping and production — exemplifies this challenge.
While AETI delivers high-performance embedded solutions to industrial clients,
the lifecycle management of deployed hardware nodes once they leave the
laboratory remains a manually intensive process. Remote deployment of new
workloads, configuration updates, and firmware changes to field-installed boards
requires on-site intervention that is difficult to scale and represents a
significant operational cost. The need for a cloud-native orchestration layer
capable of managing deployed embedded nodes autonomously and remotely defines
the industrial motivation of this work.

% ------------------------------------------------------------------
%  PARAGRAPH 3 — Problematic: three unresolved technical gaps
% ------------------------------------------------------------------
    Addressing this need through existing technologies reveals three
simultaneously unresolved technical challenges. First, although cloud-native
container orchestration frameworks such as \textit{Kubernetes} offer powerful
declarative management capabilities, their standard distributions impose a
control-plane overhead that is incompatible with resource-constrained IoT edge
hardware~\cite{bohm2022, ogbuachi2020}. Second, field-deployed nodes typically
operate behind Network Address Translation (\textit{NAT}) barriers and dynamic
IP assignments, preventing a centralised control plane from establishing reliable
communication with them without a dedicated secure tunnelling solution. Third,
the initial onboarding of an edge node into a managed cluster requires credential
distribution and network configuration steps that, in the current state of the
art, depend on physical proximity or manual operator intervention~\cite{boskov2020}.
No published architecture resolves all three constraints within a single,
integrated system.

% ------------------------------------------------------------------
%  PARAGRAPH 4 — Objectives and proposed approach
% ------------------------------------------------------------------
    This thesis proposes, implements, and experimentally validates an integrated
architecture for Zero-Touch Deployment (\textit{ZTD}) and cloud-native
orchestration of IoT edge nodes, designed to resolve these three challenges
simultaneously. The proposed system combines a lightweight \textit{K3s} cluster
hosted on \textit{AWS EC2} as the cloud control plane, an \textit{ArgoCD}
GitOps engine enforcing declarative, repository-driven deployments, and a
\textit{WireGuard}-based encrypted overlay network governed by a novel
Tri-State Fallback algorithm that ensures robust NAT traversal across all
reachable network topologies~\cite{donenfeld2017, wireguard2024}. A custom
Python Zero-Touch Provisioning daemon, integrated with \textit{AWS Lambda}
and \textit{SSM Vault} as a cloud Rendezvous \textit{API}, enables a newly
powered edge node to autonomously discover the control plane, receive its
identity credentials, establish the secure tunnel, and register into the
cluster — all without any physical human intervention. The practical utility
of this infrastructure is demonstrated through the deployment of an IoT
application that collects environmental telemetry from downstream
\textit{ESP32} microcontrollers via the \textit{CoAP} protocol and
orchestrates Over-The-Air (\textit{OTA}) firmware updates to those devices,
with end-to-end observability provided through \textit{InfluxDB~v2} and
\textit{Grafana}.

% ------------------------------------------------------------------
%  PARAGRAPH 5 — Validation methodology
% ------------------------------------------------------------------
    The system is validated through a deliberate two-phase hardware strategy.
In the first phase, the complete provisioning and orchestration pipeline is
built and verified in a virtualised environment, enabling rapid iteration and
architectural refinement without hardware constraints. In the second phase, the
validated architecture is migrated to a physical \textit{Raspberry Pi}
operating as a bare-metal edge gateway, confirming portability across hardware
substrates and readiness for real-world industrial deployment.

% ------------------------------------------------------------------
%  PARAGRAPH 6 — Thesis plan (one sentence per chapter — ENSTICP norm)
% ------------------------------------------------------------------
    This document is organised into four chapters. \textbf{Chapter~I} surveys
the state of the art in cloud-edge orchestration, Zero-Touch Provisioning,
and secure tunnelling, and establishes the technological foundation and
comparative justification for the selected stack. \textbf{Chapter~II} presents
the global architecture and detailed design of the proposed system, including
the Tri-State Fallback algorithm and the ZTP provisioning sequence.
\textbf{Chapter~III} describes the full implementation of the system, from
cloud infrastructure provisioning and GitOps pipeline configuration to the
edge node runtime and IoT application deployment. \textbf{Chapter~IV} presents
the experimental results, validates the system against its functional
requirements, and discusses the contributions and perspectives of this work.
